Flexibility of interactions depends on habitat suitability

### 2019-09-09 paper/manuscript.Rmd
Missing paragraph break in intro
remove reference to multiple model sets

Perhaps, because it operates at scales larger than those of the Eltonian niche concept, the abiotic environment is one of the factors that is commonly overlooked by community ecologists when explaining interaction turnover [but see @gravel_bringing_2018]. 
The most evident—and best-studied—way through which the environment can affect the structure of ecological networks is by way of patterns of community composition. 
At the heart of biogeography is the idea that the abiotic environment can determine the "Grinellian" niche of a species, this is whether the species can have a positive growth-rate when rare in a particular location.
For two species to interact, they first need to co-occur in the same community, which can only happen if the environmental conditions of the community fall within the Grinellian niche of both species. 
Although less understood, the environment can also affect the interactions themselves, after community composition has been established [@tylianakis_ecological_2017]. 
For instance, some studies show that trophic interactions, like predation [@mckinnon_lower_2010; @vucic-pestic_warming_2011] and herbivory [@baskett_latitudinal_2018], can increase with temperature but might decrease with precipitation [@pires_predicted_2016], and both plant-plant [@bertness_positive_1994] and plant-pollinator [@tur_evaluating_2016] interactions can switch from facilitative to competitive along an elevation/temperature gradient. 
Contrastingly, other studies have shown either no effect or non-linear effects of temperature or precipitation on plant-pollinator interactions [@devoto_patterns_2005; @gravel_bringing_2018].
One way to summarise the seemingly contradictory evidence of whether the environment can influence whether two species interact or not is by moving from gradients of temperature or precipitation to gradients of environmental stress. 
Previous research is so far equivocal and, taken together, suggest that environmental stress can have two contrasting effects on species' Eltonian-niche [@tylianakis_ecological_2017].
On the one hand, it is possible that when species are under environmental stress, they might be "pressured" to focus on partners with which they are best adapted to interact. 
For instance, @hoiss_altitude_2012 found increased phylogenetic clustering between plants and pollinators at higher altitudes; while @peralta_phylogenetic_2015 found that parasitoids in plantation forest, where environmental stress was higher than in native forests, were constrained to interact with hosts they were best adapted to attack.
Similarly, @lavandero_genotype_2013 found that environmental stress due to higher temperature reduced the breadth of the Eltonian niche of parasitoids. 
On the other, it is also possible that when species are under environmental stress, they are forced to be more flexible in their interactions as higher environmental stress is likely to be reflected in greater energetic or reproductive costs and therefore they might not be able to sustain encounter rates with their preferred partners at sufficient levels. 
In line with this hypothesis, @hoiss_interactive_2015 found that the specialisation of plant-pollinator networks decreased both with elevation and after extreme drought events.
Likewise, @pellissier_spatial_2010 found a positive relationship between niche breadth and environmental stress: disk- or bowl-shaped blossoms (which allow a large number of potential pollinator species to access pollen and nectar rewards) dominated at high altitude flower communities.

@@

Within each model set, we evaluate four models to assess the relative importance of suitability. 
A first model, our baseline model, included four population-level predictors and two grouping levels, species and the community. 
The population-level predictors in the baseline model, commonly called fixed effects, were the habitat suitability, the species guild (whether it was a plant or a pollinator), a metric of the species generalism, and its number of known possible partners.
We included the overall level of generalism as one would expect species with a large number of interactions globally to be more likely to interact with a large number of species locally. 
Concretely, we estimated an overall species level of generalism as the total number of interacting species across communities in our dataset. 
We also included the number of known possible partners as a predictor in our models as it allows us to control for the environmental effects on species co-occurrence. 
We calculate this metric by determining the number of partners with which the species is known to interact in any other community.
Controlling for the number of potential partners makes our model a particularly stringent test of our environmental stress hypotheses. 
Often the potential and the actual number of partners is the same or very close to each other, especially for rare species present only in a few communities.  
As we were particularly interested in understanding whether the effect of habitat suitability is conditional on the species guild (plant or pollinator), we, therefore, included guild and its interaction with suitability in the model. 
We allowed the intercept of degree and slope of the suitability-degree relationship to vary among species, this allowed us to investigate two questions. 
First, it allows us to inspect the extent to which suitability is a population or a group level effect. 
Second, by investigating the correlation between the intercept and the slope as a model parameter, it allowed us to inspect the extent by which species with a small or large number of interactions respond to increasing levels of environmental stress. 
To account for unmeasured differences between communities, like sampling effort, sampling method, or diversity, we calculated an intercept for each community in our study.
To facilitate model interpretation and convergence, we scaled all continuous variables to have a mean of zero and a unit standard deviation.

We compare this baseline model with three alternative models in which we remove one predictor at a time.
To quantify the difference between models, in terms of their expected out-of-sample performance, we use the Wanatabe-Akaike information criterion (WAIC). 
All models were fitted under a bayesian framework using the R package `brms` `r as.character(packageVersion('brms'))` [@burkner_brms_2017; @burkner_advanced_2018] as an interface for Stan [@carpenter_stan_2017]. 
For each model, we used four Markov chains of 4,000 iterations each; half of the iterations were used for warmup.
We used weakly informative priors for all model parameters. 
Specifically we used normal priors of mean zero and standard deviation ten for the population-level effects and the intercepts, a half-Cauchy prior with a location of zero and a scale of two for the standard deviations, and, when applicable, an LKJ-correlation prior with parameter \(\zeta = 1\) for the correlation matrix between group-level parameters.

>>>>

Perhaps, because it operates at scales larger than those of the Eltonian niche concept, the abiotic environment is one of the factors that is commonly overlooked by community ecologists when explaining interaction turnover [but see @gravel_bringing_2018]. 
The most evident—and best-studied—way through which the environment can affect the structure of ecological networks is by way of patterns of community composition. 
At the heart of biogeography is the idea that the abiotic environment can determine the "Grinellian" niche of a species, this is whether the species can have a positive growth-rate when rare in a particular location.
For two species to interact, they first need to co-occur in the same community, which can only happen if the environmental conditions of the community fall within the Grinellian niche of both species. 
Although less understood, the environment can also affect the interactions themselves, after community composition has been established [@tylianakis_ecological_2017]. 
For instance, some studies show that trophic interactions, like predation [@mckinnon_lower_2010; @vucic-pestic_warming_2011] and herbivory [@baskett_latitudinal_2018], can increase with temperature but might decrease with precipitation [@pires_predicted_2016], and both plant-plant [@bertness_positive_1994] and plant-pollinator [@tur_evaluating_2016] interactions can switch from facilitative to competitive along an elevation/temperature gradient. 
Contrastingly, other studies have shown either no effect or non-linear effects of temperature or precipitation on plant-pollinator interactions [@devoto_patterns_2005; @gravel_bringing_2018].

One way to summarise the seemingly contradictory evidence of whether the environment can influence whether two species interact or not is by moving from gradients of temperature or precipitation to gradients of environmental stress. 
Previous research is so far equivocal and, taken together, suggest that environmental stress can have two contrasting effects on species' Eltonian-niche [@tylianakis_ecological_2017].
On the one hand, it is possible that when species are under environmental stress, they might be "pressured" to focus on partners with which they are best adapted to interact. 
For instance, @hoiss_altitude_2012 found increased phylogenetic clustering between plants and pollinators at higher altitudes; while @peralta_phylogenetic_2015 found that parasitoids in plantation forest, where environmental stress was higher than in native forests, were constrained to interact with hosts they were best adapted to attack.
Similarly, @lavandero_genotype_2013 found that environmental stress due to higher temperature reduced the breadth of the Eltonian niche of parasitoids. 
On the other, it is also possible that when species are under environmental stress, they are forced to be more flexible in their interactions as higher environmental stress is likely to be reflected in greater energetic or reproductive costs and therefore they might not be able to sustain encounter rates with their preferred partners at sufficient levels. 
In line with this hypothesis, @hoiss_interactive_2015 found that the specialisation of plant-pollinator networks decreased both with elevation and after extreme drought events.
Likewise, @pellissier_spatial_2010 found a positive relationship between niche breadth and environmental stress: disk- or bowl-shaped blossoms (which allow a large number of potential pollinator species to access pollen and nectar rewards) dominated at high altitude flower communities.

@@

We evaluate four models to assess the relative importance of suitability. 
A first model, our baseline model, included five population-level predictors and two grouping levels, species and the community. 
The population-level predictors in the baseline model, commonly called fixed effects, were the habitat suitability, the species guild (plant or a pollinator), a metric of the species generalism, and its number of known possible partners.
We included the overall level of generalism as one would expect species with a large number of interactions globally to be more likely to interact with a large number of species locally. 
Concretely, we estimated an overall species level of generalism as the total number of interacting species across communities in our dataset. 
We also included the number of known possible partners as a predictor in our models as it allows us to control for the environmental effects on species co-occurrence. 
We calculate this metric by determining the number of partners with which the species is known to interact in any other community.
Controlling for the number of potential partners makes our model a particularly stringent test of our environmental stress hypotheses. 
Often the potential and the actual number of partners is the same or very close to each other, especially for rare species present only in a few communities.  
As we were particularly interested in understanding whether the effect of habitat suitability is conditional on the species guild (plant or pollinator), we, therefore, included guild and its interaction with suitability in the model. 
We allowed the intercept of degree and slope of the suitability-degree relationship to vary among species, this allowed us to investigate two questions. 
First, it allows us to inspect the extent to which suitability is a population or a group level effect. 
Second, by investigating the correlation between the intercept and the slope as a model parameter, it allowed us to inspect the extent by which species with a small or large number of interactions respond to increasing levels of environmental stress. 
To account for unmeasured differences between communities, like sampling effort, sampling method, or diversity, we calculated an intercept for each community in our study.
To facilitate model interpretation and convergence, we scaled all continuous variables to have a mean of zero and a unit standard deviation.

We compared this baseline model with three alternative models in which we remove one predictor at a time.
To quantify the difference between models, in terms of their expected out-of-sample performance, we use the Wanatabe-Akaike information criterion (WAIC). 
All models were fitted under a bayesian framework using the R package `brms` `r as.character(packageVersion('brms'))` [@burkner_brms_2017; @burkner_advanced_2018] as an interface for Stan [@carpenter_stan_2017]. 
For each model, we used four Markov chains of 4,000 iterations each; half of the iterations were used for warmup.
We used weakly informative priors for all model parameters. 
Specifically we used normal priors of mean zero and standard deviation ten for the population-level effects and the intercepts, a half-Cauchy prior with a location of zero and a scale of two for the standard deviations, and, when applicable, an LKJ-correlation prior with parameter \(\zeta = 1\) for the correlation matrix between group-level parameters.
